### 2020-03-20 ChapterImplementation/chapterImplementation.tex
Updated: Proof-read of Implementation

\section{System Architecture}
The simulation space is partitioned into regions, with each region consisting of its own instance of PhysX running on a dedicated GPU-enabled machine in the cloud. The boundary between regions is defined as a vertical plane, two-dimensionally dividing the simulation space into one region per server. The network library RakNet is used for all message passing between servers. RakNet ensures messages exhibit best effort and are received in sent order. For the purposes of simplicity, the network tick rate (the rate at which RakNet is updated) is the same as the main update rate (the rate at which the main logic of the program is updated, such as processing user input and handling callbacks from PhysX). This removes the need for the aura calculation to account for an extra time delay between network update and main update.

@@

When objects traverse region boundaries, they are added to a migration buffer with all information required to duplicate an object at a neighbouring server. The contents of each migration buffer are sent to the associated server, which will then be responsible for hosting the migrated objects. When migration messages are received, the migrated objects are created within the receiving server's simulation.

@@

For the implementation of this project, Amazon Web Services (AWS) was selected as the cloud provider, giving access to GPU-enabled instances (G2 instances). The PhysX Samples project (a publicly available project for demonstrating the features and capabilities of PhysX) was used as a starting point for this project, as this was a relatively minimal project that uses PhysX but with some useful features for use in development of this project, such as a basic renderer and many helper methods for using PhysX. For the purposes of automated testing and for running experiments, the testing library, Gtest, was used.

\begin{figure}[t]
	\centering
	\scalebox{1.5}{\input{Figs/OurDeploymentV2.pdf_tex}}
	\caption{Network Topology. Each server is an EC2 instance in the cloud, running our implementation of AP, which includes an instance of PhysX. The client is a local machine running the visualiser built in Unreal Engine. Servers connect to each other in order to exchange messages for aura creation/update and deletion and the migration of objects. The client receives replica updates from each server as well as sending client-created replicas and static objects, and RPCs.}
	\label{SystemOverview}
\end{figure}

@@

\begin{figure}[!t]	
	\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm,scale=1.5]
	\clip(-2.25,0) rectangle (8.4,3.75);
	\fill[line width=0.8pt,color=sim,fill=sim,fill opacity=0.1] (-2.25,3) -- (-2.25,0) -- (8.25,0) -- (8.25,3) -- cycle;
	\draw (-1.5,3.5) node[anchor=center] {Server 0};
	\draw (1.5,3.5) node[anchor=center] {Server 1};
	\draw (4.5,3.5) node[anchor=center] {\textbf{...}};
	\draw (7.5,3.5) node[anchor=center] {Server N-1};
	\draw [line width=2pt] (0,0)-- (0,3);
	\draw [line width=2pt] (3,0)-- (3,3);
	\draw [line width=2pt] (6,0)-- (6,3);
	\end{tikzpicture}
	\centering
	\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm]
	\clip(-4,0) rectangle (7,1.5);
	\draw [line width=2pt] (-0.5,0.7)-- (-0.3,0.7);
	\draw (-0.1,0.7) node[anchor=west] {Server-Region Boundary};
	\end{tikzpicture}
	\caption{Column Layout}
	\label{ServerColumnLayout}
\end{figure}

@@

\begin{figure}[!t]
	\begin{subfigure}{.5\textwidth}
		\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm,scale=1.2]
		\fill[line width=0.8pt,color=sim,fill=sim,fill opacity=0.1] (-3,3) -- (-3,-2) -- (3,-2) -- (3,3) -- cycle;
		\draw [line width=2pt] (-3,0)-- (3,0);
		\draw [line width=2pt] (0,3)-- (0,0);
		\draw (-1.5,2.5) node[anchor=center] {Server 0};
		\draw (1.5,2.5) node[anchor=center] {Server 1};
		\draw (0,-1) node[anchor=center] {Server 2};
		\end{tikzpicture}
		\centering
		\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm]
		\draw [line width=2pt] (-2.5,0.5)-- (-2.3,0.5);
		\draw (-2.1,0.5) node[anchor=west] {Region Boundary};
		\end{tikzpicture}
		\caption{3 Server Corner Topology}
		\label{3ServerCornerLayout}
	\end{subfigure}%
	\begin{subfigure}{.5\textwidth}
		\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm,scale=1.2]
		%	\clip(-3,-3) rectangle (3,3);
		\fill[line width=0.8pt,color=sim,fill=sim,fill opacity=0.1] (-3,3) -- (-3,-3) -- (3,-3) -- (3,3) -- cycle;
		\draw [line width=2pt] (-3,0)-- (3,0);
		\draw [line width=2pt] (0,3)-- (0,0);
		\draw (-1.5,2.5) node[anchor=center] {Server 0};
		\draw (1.5,2.5) node[anchor=center] {Server 1};
		\draw (-1.5,-2.5) node[anchor=center] {Server 2};
		\draw (1.5,-2.5) node[anchor=center] {Server 3};
		\draw [line width=2pt] (0,0)-- (0,-3);
		\end{tikzpicture}
		\centering
		\caption{4 Server Corner Topology}
		\label{4ServerCornerLayout}
	\end{subfigure}
	\caption{Corner Layouts}
	\label{ServerCornerLayout}
\end{figure}

@@

For each boundary with a neighbouring server, a server has three message buffers 1) A migration buffer for sending objects to the server corresponding with that boundary. Each element in the buffer is a message for one object migration. Each migration message contains all data necessary to reproduce the object on the receiving server. 2) An aura creation/update buffer, each element of the buffer is a message for the creation or update of an aura and includes ID, position and radius of the aura. The contents of this buffer are sent to all neighbouring servers, not just the server corresponding with this boundary. The messages structure for both creation and update are identical as the receiving server creates a new aura if one does not already exist for the corresponding object (determined using the ID), otherwise the message is just used to update the state of the existing aura. 3) An aura delete buffer, each element of the buffer is a message for the deletion of an aura, which consists of the aura's ID. The contents of this buffer are sent to all neighbouring servers, not just the server corresponding with this boundary.

@@

\subsection{Future Architecture}
System architectures which could be investigated in the future will now be proposed and discussed. A limitation of the current architecture is that a client must connect to all servers; servers that are responsible for running the simulation also have the overhead of connecting to many clients. An alternative to this is to make the use of an `overseer' node, which acts as a middle-man for communication with all the servers, as illustrated in Fig. \ref{OverseerNode}. The overseer node is co-located in the cloud with the simulation servers. The clients all connect to the overseer node, but all the servers only connect to one overseer node, reducing the overhead of many clients connecting to a server, freeing up computational resources on each simulation server, allowing more resources to be used for the simulation itself. An additional advantage of this approach is it allows for some fault-tolerance. If one or more simulation servers were to fail, the overseer node has a copy of the states of all objects from those servers, meaning those spatial partitions of the simulation could be restored. Another advantage of overseers is it can potentially allow  some computational work to be carried out on the overseer nodes instead of the simulation nodes, such as simulation queries and non-simulation work, such as game logic, interest management.
The disadvantage of using an overseer node is the additional latency between client and simulation server, messages now have to be processed by the overseer node between the simulation servers and clients. For example, if a replicated object moves in one of the simulation servers, the update message for that object has to be processed and sent by the simulation server, sent over the network to the overseer, received and processed by the overseer node and then sent over the network to all connected clients. The additional latency is due to the communication latency between simulation server and overseer node and the time taken to process messages on the overseer node.

@@

\section{Aura Implementation}
How the auras in AP are implemented will now be described at a high-level, before a formal description of the algorithms using pseudo-code is given in \ref{Algorithms}.
Auras are implemented through the use of trigger volumes (non-physically interacting volumes, that generate trigger collision callbacks). Two trigger volumes are used for different purposes. 
The first is used for auras that are being received from remote servers. When an aura is received, a trigger volume, consisting of a sphere, is created at the corresponding position with a given radius. The second trigger volume used is by the server-region boundaries. Each server-region boundary has a trigger volume that extends inside the server's region, it extends the same distance as the aura radius, minus the bounding sphere of an object. This is used to determine when an object should begin and stop having its aura sent to neighbouring servers and avoid the need for per-object checks for proximity to the server-region boundary. When an object collides with the aura trigger boundary, the aura begins to be sent, when an object leaves the aura trigger boundary, an aura delete message is sent.
Each server-region boundary has a trigger volume, objects may collide with multiple aura trigger boundaries, meaning auras can be sent to multiple servers, as illustrated in Fig. \ref{AuraCorner}.

@@

When a server begins communicating an aura, a message is added to the aura creation/update buffer. This message consists of the position of the centre of the aura, the radius of the aura. % and a flag signifying whether or not this is a creation message or not. Flag is not currently used as often causes issue in the corner case, possibly due to multiple types of messages concerncing one object being sent and recieved in the same tick.
The contents of the buffer are then broadcast to all neighbouring servers. When aura creation/update messages are received and an aura for that object does not already exist, a new aura, a sphere trigger volume, is created at that position with the given radius.
Whenever an object projecting an aura moves, and update message is added to the aura creation/update buffer, containing the new position and radius of the aura (aura radius is currently fixed, but may become dynamic in the future). The contents of the buffer are broadcast to all neighbouring servers. When the message is received, the positions of the auras are updated. 
If an object leaves aura trigger boundary, either by moving away from the boundary or by being migrated to another server, a delete aura message is added to the aura delete buffer. The contents of the buffer are then broadcast to all neighbouring servers. When the message is received, the aura trigger volume is removed from the simulation.

@@

In addition to auras existing as trigger volumes representing auras of remote objects, servers also track the auras of objects that they are hosting, that are having their auras broadcast. These are used to determine clusters of objects, so that clusters of objects are migrated together, at the same time. When an object begins broadcasting an aura, a query-only shape is attached to the object, which represents the object's aura. When an object is a migration candidate, a scene query is carried out, using this query-only shape, to determine which auras overlap the migration candidate's aura. Using a recursive search, involving a scene query for each object found from the prior query, it is possible to determine the objects in the cluster. If the entire cluster is outside of the host server's region, the entire cluster is added to the migration buffer and sent to the server in which the cluster resides. If any object in the cluster is still within the host server's region, no migration takes place. This prevents clusters being repeatedly being migrated back and forth between servers.

@@

\section{Algorithms}\label{Algorithms}
$AC$ is called when an object collides with an aura. A recursive search is performed in order to find all objects that would lie within each object's aura, preventing thrashing as discussed in \ref{Thrashing}. Once the recursive search is complete, all objects are added to the send buffer.

@@

$BT$ is called when an object traverses a boundary. In order to prevent `islands' forming (for example Objects H, I and J in Fig. \ref{AuraProj}), a recursive search is carried out to determine if an object is part of an island or not. If an object is found to not be part of an island, the entire cluster of objects is added to the send buffer, otherwise no action is taken.

@@

From the Boundary Traversed state, an object remains in this state until one of the following occurs: the object begins intersecting the host's boundary again (triggered through a PhysX trigger collision callback), in which case the object moves to the Intersecting Boundary state; the object collides with a remote object's aura (triggered through a PhysX callback, the AC algorithm is run), in which case it is immediately moved into the Sent state; or the object cluster (group of one or more object's with overlapping auras) becomes an island (no object in cluster is intersecting the host's boundary) determined using the BT algorithm, in which case the object is moved to the sent state. The object cluster island check (BT algorithm) is performed every update, per boundary.

@@

\subsection{Overview}
The visualiser connects to multiple servers, objects are simulated on server and those objects from all servers are replicated on the visualiser. These are referred to as replica objects. The visualiser uses RakNet's ReplicaManager, which is responsible for: the creation of replicas objects on both the server and client; communicating the state of objects from the server to the client, in order for the replica objects to be replicated; and the deletion of replica objects. Replica objects are replicated on the visualiser by the server sending the state (position, rotation) to the client, the replica representation on the client is then updated with the newly received state from the server. The object state is optimised to reduce processing and network overhead, by only sending updates of the position and rotation to the client, when they change on the server. In the visualiser, replica objects are kinematic objects, i.e. they are not physically simulated in the visualiser, their physical states are updated entirely by messages received from the server, such as their movement from velocity and no interactions, such as collisions are simulated in the visualiser.

@@

\subsection{Server-created replica objects}\label{Server-CreatedReplicas}
The visualiser supports server-created replica objects, which are limited to pre-defined basic objects (cuboid, sphere \& capsule). A different render material per server is used to identify the owning server of each object. These server objects have pre-defined representations on the client. The server sends an identifier for the type of object, which is then created on the client. Using pre-defined objects means the whole geometry of the object does not need to be transferred to the client. This is useful for when there are a large number of identical objects as it avoids having to re-send the same geometry for each object. However, this method requires the client to know of these pre-defined types.

@@

\subsection{Client-created static objects}\label{Client-CreatedReplicas}
In addition to replica objects, static objects can also be created by the client. As these objects are static, there is no need for their states to be replicated as they do not change over the course of the simulation. Static object work simply by the client sending the geometry, position and rotation details of the static object to all servers and the servers then re-create the static object. These static objects then remain unchanged throughout the simulation, requiring no updates to the state to be communicated to the client. Static objects are created on all servers, this could be optimised so only static objects within a server's region are created, however, AP can lead to clusters of objects being simulated outside of their host server's region, which may interact with static objects that do not exist on their host server. In addition, the cost of simulating static objects tends to be very low. They do not require any updates as they are not dynamic. In the future, region boundaries may change; a server may become responsible for a static object previously residing in another server's region, which would then have to be communicated when the region boundaries changed, by all servers simulating all static objects, this avoids the need for extra communication when server boundaries change. With these trade-offs in mind, we have elected to create all static objects on all servers, when static objects are used in a simulation.

\subsection{Interactivity}\label{Interactivity}
The visualiser also allows for interactivity, the creation of server-created objects can be triggered from the client. This feature uses RakNet's Remote Procedure Calls (RPCs), which enables clients to execute procedures on the server, including any required parameters. The client sends a RPC (triggered by a key press) with the parameters of player position and forward direction. The RPC message is received by all servers, only the server which is responsible for the player position creates a server-created object at that position and the object is given a velocity in the direction of the player's direction. This allows the player to create new objects at run-time, which are then able to interact with all objects in the simulation.

@@

\subsection{Future work}
Currently when server created objects are migrated between servers, the client destroys the replica and a new replica is created. This is because the simulated object is destroyed on one server and re-created on another. A better way of dealing with this, would be a procedure for handing over server-ownership of replicas, removing the need for replicas to be destroyed and recreated. This would reduce processing overhead on the client, as there is no longer a need to destroy and recreate objects when server migrations take place.

@@

Currently all objects from all servers are replicated on the client with equal priority and frequency. Interest management would be required for large complex scenes as the visualiser can only replicate and render a limited number of objects.

Mesh deformation would require significant network traffic, especially if done over a period of time, rather than in a single instance. Position and rotation can be replicated with small amounts of data, however, geometry data can be arbitrarily large. In mesh deformations 100s or even 1000s of vertex points can change, all of which would have to be serialised and transmitted over the network in order to be correctly replicated on the client.

>>>>

\section{System Architecture}
The simulation space is partitioned into regions, with each region consisting of its own instance of PhysX running on a dedicated GPU-enabled machine in the cloud. The boundary between regions are defined as vertical planes, two-dimensionally dividing the simulation space into one region per server. The network library RakNet is used for all message passing between servers. RakNet ensures messages exhibit best effort and are received in sent order. For the purposes of simplicity, the network tick rate (the rate at which RakNet is updated) is the same as the main update rate (the rate at which the main logic of the program is updated, such as processing user input and handling callbacks from PhysX). This removes the need for the aura calculation to account for an extra time delay between network update and main update.

@@

When objects fully traverse region boundaries, they are added to a migration buffer with all information required to duplicate an object at a neighbouring server. The contents of each migration buffer are sent to the associated server, which will then be responsible for hosting the migrated objects. When migration messages are received, the migrated objects are created within the receiving server's simulation.

@@

For the implementation of this project, Amazon Web Services (AWS) was selected as the cloud provider, giving access to GPU-enabled instances (G2 instances). PhysX makes us of GPU-acceleration, including for the simulation of rigid-bodies \cite{GPURigid}. The PhysX Samples project (a publicly available project for demonstrating the features and capabilities of PhysX) was used as a starting point for this project, as this was a relatively minimal project that uses PhysX but with some useful features for use in development of this project, such as a basic renderer and many helper methods for using PhysX. For the purposes of automated testing and for running experiments, the testing library, Gtest, was used.

\begin{figure}[t]
	\centering
	\scalebox{1.5}{\input{Figs/OurDeploymentV2.pdf_tex}}
	\caption{Network Topology. Each server is an EC2 instance in the cloud, running AP, which includes an instance of PhysX. The client is a local machine running the visualiser built in Unreal Engine. Servers connect to each other in order to exchange messages for aura creation/update and deletion and the migration of objects. The client receives replica updates from each server as well as sending client-created replicas, static objects and RPCs.}
	\label{SystemOverview}
\end{figure}

@@

\begin{figure}[!t]	
	\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm,scale=1.5]
	\clip(-2.25,0) rectangle (8.4,3.75);
	\fill[line width=0.8pt,color=sim,fill=sim,fill opacity=0.1] (-2.25,3) -- (-2.25,0) -- (8.25,0) -- (8.25,3) -- cycle;
	\draw (-1.5,3.5) node[anchor=center] {Server 0};
	\draw (1.5,3.5) node[anchor=center] {Server 1};
	\draw (4.5,3.5) node[anchor=center] {\textbf{...}};
	\draw (7.5,3.5) node[anchor=center] {Server N-1};
	\draw [line width=2pt] (0,0)-- (0,3);
	\draw [line width=2pt] (3,0)-- (3,3);
	\draw [line width=2pt] (6,0)-- (6,3);
	\end{tikzpicture}
	\centering
	\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm]
	\clip(-4,0) rectangle (7,1.5);
	\draw [line width=2pt] (-0.5,0.7)-- (-0.3,0.7);
	\draw (-0.1,0.7) node[anchor=west] {Server-Region Boundary};
	\end{tikzpicture}
	\caption{Column topology of server regions}
	\label{ServerColumnLayout}
\end{figure}

@@

\begin{figure}[!t]
	\begin{subfigure}{.5\textwidth}
		\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm,scale=1.2]
		\fill[line width=0.8pt,color=sim,fill=sim,fill opacity=0.1] (-3,3) -- (-3,-2) -- (3,-2) -- (3,3) -- cycle;
		\draw [line width=2pt] (-3,0)-- (3,0);
		\draw [line width=2pt] (0,3)-- (0,0);
		\draw (-1.5,2.5) node[anchor=center] {Server 0};
		\draw (1.5,2.5) node[anchor=center] {Server 1};
		\draw (0,-1) node[anchor=center] {Server 2};
		\end{tikzpicture}
		\centering
		\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm]
		\draw [line width=2pt] (-2.5,0.5)-- (-2.3,0.5);
		\draw (-2.1,0.5) node[anchor=west] {Region Boundary};
		\end{tikzpicture}
		\caption{3 server corner topology}
		\label{3ServerCornerLayout}
	\end{subfigure}%
	\begin{subfigure}{.5\textwidth}
		\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm,scale=1.2]
		%	\clip(-3,-3) rectangle (3,3);
		\fill[line width=0.8pt,color=sim,fill=sim,fill opacity=0.1] (-3,3) -- (-3,-3) -- (3,-3) -- (3,3) -- cycle;
		\draw [line width=2pt] (-3,0)-- (3,0);
		\draw [line width=2pt] (0,3)-- (0,0);
		\draw (-1.5,2.5) node[anchor=center] {Server 0};
		\draw (1.5,2.5) node[anchor=center] {Server 1};
		\draw (-1.5,-2.5) node[anchor=center] {Server 2};
		\draw (1.5,-2.5) node[anchor=center] {Server 3};
		\draw [line width=2pt] (0,0)-- (0,-3);
		\end{tikzpicture}
		\centering
		\caption{4 server corner topology}
		\label{4ServerCornerLayout}
	\end{subfigure}
	\caption{Corner topologies of server regions}
	\label{ServerCornerLayout}
\end{figure}

@@

For each boundary with a neighbouring server, a server has three message buffers:
\begin{enumerate}
	\item A migration buffer for sending objects to the server corresponding with that boundary. Each element in the buffer is a message for one object migration. Each migration message contains all data necessary to reproduce the object on the receiving server. 
	\item An aura creation/update buffer, each element of the buffer is a message for the creation or update of an aura and includes ID, position and radius of the aura. The contents of this buffer are sent to all neighbouring servers, not just the server corresponding with this boundary. The messages structure for both creation and update are identical as the receiving server creates a new aura if one does not already exist for the corresponding object (determined using the ID), otherwise the message is just used to update the state of the existing aura. 
	\item An aura delete buffer, each element of the buffer is a message for the deletion of an aura, which consists of the aura's ID. The contents of this buffer are sent to all neighbouring servers, not just the server corresponding with this boundary.
\end{enumerate}

@@

\subsection{Future Architecture}
System architectures which could be investigated in the future will now be proposed and discussed. A limitation of the current architecture is that a client must connect to all servers; servers that are responsible for running the simulation also have the overhead of connecting to many clients. An alternative to this is to make the use of an `overseer' node, which acts as a middle-man for communication with all the servers, as illustrated in Fig. \ref{OverseerNode}. The overseer node is co-located in the cloud with the simulation servers. The clients all connect to the overseer node, but all the servers only connect to one overseer node, reducing the overhead of many clients connecting to a server, freeing up computational resources on each simulation server, allowing more resources to be used for the simulation itself. An additional advantage of this approach is it allows for some fault-tolerance. If one or more simulation servers were to fail, the overseer node has a copy of the states of all objects from those servers, meaning those spatial partitions of the simulation could be restored. Another advantage of overseers is it can potentially allow some computational work to be carried out on the overseer nodes instead of the simulation nodes, such as simulation queries and non-simulation work, such as game logic, interest management.

The disadvantage of using an overseer node is the additional latency between client and simulation server, messages now have to be processed by the overseer node between the simulation servers and clients. For example, if a replicated object moves in one of the simulation servers, the update message for that object has to be processed and sent by the simulation server, sent over the network to the overseer, received and processed by the overseer node and then sent over the network to all connected clients. The additional latency is due to the communication latency between simulation server and overseer node and the time taken to process messages on the overseer node.

@@

\section{Aura Implementation}
How the auras in AP are implemented will now be described at a high-level, before a formal description of the algorithms using pseudo-code are given in \ref{Algorithms}.
Auras are implemented through the use of trigger volumes (non-physically interacting volumes, that generate trigger collision callbacks). Two trigger volumes are used for different purposes. 
The first is used for auras that are being received from remote servers. When an aura is received, a trigger volume, consisting of a sphere, is created at the corresponding position with a given radius. The second trigger volume used is by the server-region boundaries. Each server-region boundary has a trigger volume that extends inside the server's region, it extends the same distance as the aura radius, minus the bounding sphere of an object. This is used to determine when an object should begin and stop having its aura sent to neighbouring servers and avoid the need for per-object checks for proximity to the server-region boundary. When an object collides with the aura trigger boundary, the aura begins to be sent, when an object leaves the aura trigger boundary, an aura delete message is sent.
Each server-region boundary has a trigger volume, objects may collide with multiple aura trigger boundaries, meaning auras can be sent to multiple servers, as illustrated in Fig. \ref{AuraCorner}.

@@

When a server begins communicating an aura, a message is added to the aura creation/update buffer. This message consists of the position of the centre of the aura and the radius of the aura. % and a flag signifying whether or not this is a creation message or not. Flag is not currently used as often causes issue in the corner case, possibly due to multiple types of messages concerncing one object being sent and recieved in the same tick.
The contents of the buffer are then broadcast to all neighbouring servers. When aura creation/update messages are received and an aura for that object does not already exist, a new aura, a sphere trigger volume, is created at that position with the given radius.
Whenever an object projecting an aura moves, and update message is added to the aura creation/update buffer, containing the new position and radius of the aura (aura radius is currently fixed, but may become dynamic in the future). The contents of the buffer are broadcast to all neighbouring servers. When the message is received, the positions of the auras are updated. 
If an object leaves aura trigger boundary, either by moving away from the boundary or by being migrated to another server, a delete aura message is added to the aura delete buffer. The contents of the buffer are then broadcast to all neighbouring servers. When the message is received, the aura trigger volume is removed from the simulation.

@@

In addition to auras existing as trigger volumes representing auras of remote objects, servers also track the auras of objects that they are hosting, that are having their auras broadcast. These are used to determine clusters of objects, so that clusters of objects are migrated together, at the same time. When an object begins broadcasting an aura, a query-only shape is attached to the object, which represents the object's aura. When an object is a migration candidate, a scene query is carried out, using this query-only shape, to determine which auras overlap the migration candidate's aura. Using a recursive search, involving a scene query for each object found from the prior query, it is possible to determine the objects in the cluster. If the entire cluster is outside of the host server's region, the entire cluster is added to the migration buffer and sent to the server in which the cluster resides. If any object in the cluster is still within the host server's region, no migration takes place. This prevents clusters being repeatedly migrated back and forth between servers.

@@

\section{Algorithms}\label{Algorithms}
In this section the algorithms used by AP will be discussed in detail and pseudocode presented. AP can be broken down into 6 main algorithms:
\begin{itemize}
	\item $AC$ - Object migration as a result of an aura collision.
	\item $BT$ - Object migration as a result of boundary traversal.
	\item $OBC$ - Aura creation, when an object collides with a server boundary.
	\item $OBU$ - Aura update, which occurs every update for all existing auras.
	\item $OBE$ - Aura destruction, when an object exits a server boundary.
	\item $BNU$ - Boundary network update, occurs every update for each network connection between servers and exchanges object migrations and aura messages for all boundaries.
\end{itemize}

$AC$ is called when an object collides with an aura. A recursive search is performed in order to find all objects that would lie within each object's aura, preventing thrashing as discussed in \ref{Thrashing}. Once the recursive search is complete, all objects are added to the send buffer.

@@

$BT$ is called when an object traverses a boundary. In order to prevent `islands' forming (for example Objects H, I and J in Fig. \ref{AuraProj}), a recursive search is carried out to determine if an object is part of an island or not. If an object is found to not be part of an island, the entire cluster of objects is added to the send buffer, otherwise no action is taken. Note that a cluster may consist of only one object, meaning the algorithm handles migration of both single and groups of objects.

@@

From the Boundary Traversed state, an object remains in this state until one of the following occurs: the object begins intersecting the host's boundary again (triggered through a PhysX trigger collision callback), in which case the object moves to the Intersecting Boundary state; the object collides with a remote object's aura (triggered through a PhysX callback, the AC algorithm is run), in which case it is immediately moved into the Sent state; or the object cluster (group of one or more objects with overlapping auras) becomes an island (no object in the cluster is intersecting the host's boundary) determined using the BT algorithm, in which case the object is moved to the sent state. The object cluster island check (BT algorithm) is performed every update, per boundary.

@@

\subsection{Overview}
The visualiser connects to multiple servers, objects are simulated on server and those objects from all servers are replicated on the visualiser. These are referred to as replica objects. The visualiser uses RakNet's ReplicaManager, which is responsible for: the creation of replicas objects on both the server and client; communicating the state of objects from the server to the client, in order for the replica objects to be replicated; and the deletion of replica objects. Replica objects are replicated on the visualiser by the server sending the state (position, rotation) to the client, the replica representation on the client is then updated with the newly received state from the server. The object state is optimised to reduce processing and network overhead, by only sending updates of the position and rotation to the client, when they change on the server. In the visualiser, replica objects are kinematic objects, i.e. they are not physically simulated in the visualiser, their physical states are updated entirely by messages received from the server, such as their movement from velocity and no interactions, such as collisions, are simulated in the visualiser.

@@

\subsection{Server-created replica objects}\label{Server-CreatedReplicas}
The visualiser supports server-created replica objects, which are limited to pre-defined basic objects (cuboid, sphere \& capsule). A different render material per server is used to identify the owning server of each object. These server objects have pre-defined representations on the client. The server sends an identifier for the type of object, which is then created on the client. Using pre-defined objects means the whole geometry of the object does not need to be transferred to the client. This is useful for when there are a large number of identical objects as it avoids having to re-send the same geometry for each object. However, this method requires the client to know of these pre-defined types before packaging.

@@

\subsection{Client-created static objects}\label{Client-CreatedReplicas}
In addition to replica objects, static objects can also be created by the client. As these objects are static, there is no need for their states to be replicated as they do not change over the course of the simulation. Static object work simply by the client sending the geometry, position and rotation details of the static object to all servers and the servers then re-create the static object. These static objects then remain unchanged throughout the simulation, requiring no updates to the state to be communicated to the client. Static objects are created on all servers, this could be optimised so only static objects within a server's region are created, however, AP can lead to clusters of objects being simulated outside of their host server's region, which may interact with static objects that do not exist on their host server. In addition, the cost of simulating static objects tends to be very low. They do not require any updates as they are not dynamic. In the future, region boundaries may change; a server may become responsible for a static object previously residing in another server's region, which would then have to be communicated when the region boundaries changed, by all servers simulating all static objects, this avoids the need for extra communication when server boundaries change. With these trade-offs in mind, the decision for AP was made to create all static objects on all servers, when static objects are used in a simulation.

\subsection{Interactivity}\label{Interactivity}
The visualiser also allows for interactivity, the creation of server-created objects can be triggered from the client. This feature uses RakNet's Remote Procedure Calls (RPCs), which enables clients to execute procedures on the server, including any required parameters. The client sends a RPC (triggered by a key press) with the parameters of player position and forward direction. The RPC message is received by all servers, only the server which is responsible for the player position creates a server-created object at that position and the object is given a velocity in the direction of the player's forward direction. This allows the player to create new objects at run-time, which are then able to interact with all objects in the simulation.

@@

\subsection{Future work}
Currently when server created objects are migrated between servers, the client destroys the replica and a new replica is created. This is because the simulated object is destroyed on one server and re-created on another. A more efficient way of dealing with this, would be a procedure for handing over server-ownership of replicas, removing the need for replicas to be destroyed and recreated. This would reduce processing overhead on the client, as there is no longer a need to destroy and recreate objects when server migrations take place. In addition, this would also prevent any chance of objects appearing to `flicker' due to any delays between being destroyed and recreated.

@@

Currently all objects from all servers are replicated on the client with equal priority and frequency. Interest management could be employed for large complex scenes as the visualiser can only replicate and render a limited number of objects.

Mesh deformation is a technique sometimes used in video games, in which the geometric mesh representation of an object is deformed, often as a result of physical interaction. Mesh deformation is not currently supported in AP and would require significant network traffic, especially if done over a period of time (a series of many deformations), rather than in a single instance. Position and rotation can be replicated with small amounts of data, however, geometry data can be arbitrarily large. In mesh deformations 100s or even 1000s of vertex points can change, all of which would have to be serialised and transmitted over the network in order to be correctly replicated on the client.
